---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    embed-resources: true
    fig-cap-location: bottom
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
params:
  title: "RNA-seq Analysis Report"
  author: ""
  outdir: "."
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
outdir <- params$outdir
deg_dir <- file.path(outdir, "1-DEG")
gsea_dir <- file.path(outdir, "2-GSEA")
viz_dir <- file.path(outdir, "3-Visualization")

read_csv_safe <- function(path, ...) {
  if (!file.exists(path)) return(NULL)
  tryCatch(utils::read.csv(path, stringsAsFactors = FALSE, ...), error = function(e) NULL)
}
list_pdfs <- function(dir, pattern = "\\.pdf$") {
  if (!dir.exists(dir)) return(character(0))
  sort(list.files(dir, pattern = pattern, full.names = FALSE, recursive = TRUE))
}
show_table <- function(df, caption = NULL, max_rows = 50) {
  if (is.null(df) || nrow(df) == 0) {
    cat("_No data available._\n")
    return(invisible(NULL))
  }
  if (nrow(df) > max_rows) {
    cat(sprintf("_Showing first %d of %d rows._\n", max_rows, nrow(df)))
    df <- utils::head(df, max_rows)
  }
  knitr::kable(df, caption = caption, row.names = FALSE)
}
show_pdf <- function(rel_path, width = "100%") {
  # Render the first PDF page to PNG before embedding. Browser PDF objects are
  # inconsistent and can turn an otherwise valid figure into a blank panel.
  full <- file.path(viz_dir, rel_path)
  if (!file.exists(full)) {
    cat("_Figure not found: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  if (!nzchar(Sys.which("pdftoppm"))) {
    cat("_PDF preview unavailable because `pdftoppm` is not installed: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  base <- tempfile(pattern = "rnaseq-report-")
  png <- paste0(base, ".png")
  code <- suppressWarnings(system2(
    "pdftoppm",
    c("-f", "1", "-l", "1", "-singlefile", "-r", "150", "-png",
      shQuote(full), shQuote(base)),
    stdout = FALSE, stderr = FALSE
  ))
  if (!identical(code, 0L) || !file.exists(png)) {
    cat("_PDF preview failed: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  uri <- knitr::image_uri(png)
  cat(sprintf(
    '<figure><img src="%s" alt="%s" style="width:%s;height:auto"><figcaption>%s</figcaption></figure>',
    uri, rel_path, width, rel_path
  ))
}

coverage <- read_csv_safe(file.path(outdir, "report_coverage_manifest.csv"))
claim_ledger <- read_csv_safe(file.path(outdir, "claim_evidence_ledger.csv"), check.names = FALSE)
inventory <- if (dir.exists(outdir)) {
  paths <- list.files(outdir, recursive = TRUE, full.names = FALSE)
  data.frame(path = paths, extension = tolower(tools::file_ext(paths)), stringsAsFactors = FALSE)
} else NULL
```

<style>
.summary-box {background:#f3f7fb;border-left:5px solid #3778a8;padding:0.9rem 1.1rem;margin:1rem 0;}
.caution-box {background:#fff8e8;border-left:5px solid #d99a2b;padding:0.9rem 1.1rem;margin:1rem 0;}
figure {margin:1.2rem 0 1.6rem 0;}
figure img {border:1px solid #e5e7eb;}
figcaption {color:#5f6368;font-size:0.9rem;margin-top:0.45rem;}
details {margin:0.8rem 0;}
</style>

# Technical summary

<div class="summary-box">

This report summarizes the saved outputs of an **RNAseq-Templates** run. It is
designed as an evidence-linked technical report: primary results appear in the
main reading path, sensitivity and exhaustive output indexes appear later, and
missing analysis domains remain visible in the coverage table.

Project-specific biological mechanisms are reported only when a validated
`claim_evidence_ledger.csv` is supplied. In its absence, this report describes
the observed statistical results without inventing a mechanism narrative.

</div>

**Analysis directory:** `r outdir`

```{r summary}
summary_txt <- file.path(outdir, "Analysis_summary.txt")
if (file.exists(summary_txt)) {
  cat("```\n")
  cat(readLines(summary_txt, warn = FALSE), sep = "\n")
  cat("\n```\n")
}
```

# Scope, samples, and inferential definitions

The experimental unit, inclusion decisions, group sizes, design formula,
contrasts, primary threshold and multiple-testing families should be defined
before interpreting differential expression or enrichment. The saved metadata
below is the report's source of truth for sample inclusion.

```{r sample-design}
coldata <- read_csv_safe(file.path(deg_dir, "colData.csv"), check.names = FALSE)
show_table(coldata, caption = "Samples included in the analysis and their modeled groups.", max_rows = 100)
```

# Sample quality supports—or limits—the modeled comparisons

QC figures are used to identify gross data-quality problems, unexpected sample
structure and possible outliers. Visual separation is not evidence of a
mechanism, and samples should not be removed simply because they weaken the
expected result.

```{r qc, results='asis'}
qc_pdfs <- list_pdfs(viz_dir, pattern = "^(PCA_plot|Sample_distance_heatmap|Sample_correlation|Raw_count_distribution|Filter_retention|Sample_QC_metrics).*\\.pdf$")
for (f in utils::head(qc_pdfs, 8)) {
  cat("\n## ", sub("\\.pdf$", "", basename(f)), "\n\n")
  show_pdf(f)
  cat("\nRead this figure at the sample level; any exclusion requires a predeclared QC rule or an independently documented technical failure.\n")
}
if (length(qc_pdfs) == 0) cat("_No QC figures found._\n")
```

# Differential expression is reported at one primary threshold

## DEG counts by threshold

```{r deg-summary}
deg_summary <- read_csv_safe(file.path(deg_dir, "DEG_threshold_summary.csv"))
show_table(deg_summary, caption = "Number of differentially expressed genes per comparison and threshold.")
```

The declared primary threshold should drive headline counts. Other rows are
sensitivity analyses and must not be searched post hoc for the most favorable
result. DEG counts depend on effect magnitude, dispersion, expression level and
sample size; they are not a standalone measure of treatment strength.

## Raw vs shrunken log2FC strategy

```{r lfc-strategy}
lfc_strategy <- read_csv_safe(file.path(deg_dir, "DEG_lfc_strategy_summary.csv"))
show_table(lfc_strategy, caption = "DEG counts using raw vs shrunken log2 fold-change.")
```

## Volcano plots

```{r volcano, results='asis'}
volcano_pdfs <- list_pdfs(viz_dir, pattern = "^Volcano_.*\\.pdf$")
if (length(volcano_pdfs) == 0) cat("_No volcano plots found._\n")
for (f in volcano_pdfs) {
  cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
}
```

## Expression heatmaps and selected genes

```{r heatmaps, results='asis'}
hm_pdfs <- list_pdfs(viz_dir, pattern = "(Heatmap_topDEG|Heatmap_key_genes|DEG_barplot_).*\\.pdf$")
if (length(hm_pdfs) == 0) cat("_No DEG or key-gene heatmaps/plots found._\n")
for (f in utils::head(hm_pdfs, 12)) {
  cat("\n### ", sub("\\.pdf$", "", basename(f)), "\n\n")
  show_pdf(f)
  cat("\nSelected-gene views support candidate inspection but do not create an independent statistical test.\n")
}
```

# ORA findings depend on the DEG threshold and tested background

```{r ora-summary}
ora_summary <- read_csv_safe(file.path(gsea_dir, "ORA_threshold_summary.csv"))
show_table(ora_summary, caption = "Over-representation analysis summary across thresholds.")
```

ORA asks whether a thresholded DEG list contains more genes from a term than
expected against the declared expressed-gene background. Direction, threshold,
ontology and correction family must be retained when interpreting a term.

```{r ora-figures, results='asis'}
ora_pdfs <- list_pdfs(viz_dir, pattern = "(ORA|GO_|KEGG_).*\\.pdf$")
if (length(ora_pdfs) > 0) {
  cat("\n## Representative ORA evidence\n\n")
  preferred <- ora_pdfs[grepl("(^|/)standard/", ora_pdfs)]
  if (!length(preferred)) preferred <- ora_pdfs
  for (f in utils::head(preferred, 10)) {
    cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
    cat("\nInterpret the displayed term together with its gene count, direction, adjusted P value and threshold-specific source table.\n")
  }
} else {
  cat("_No ORA figures found or ORA was not applicable._\n")
}
```

# GSEA provides rank-based program evidence, not literal pathway function

GSEA uses a full ranked gene list and therefore reduces dependence on a single
DEG cutoff. GO terms can be redundant and disease-labelled KEGG terms often
reflect shared ribosomal, mitochondrial or inflammatory genes. Leading-edge
genes—not the pathway title alone—should anchor biological interpretation.

```{r gsea-figures, results='asis'}
gsea_pdfs <- list_pdfs(viz_dir, pattern = "GSEA.*\\.pdf$")
if (length(gsea_pdfs) == 0) {
  cat("_No GSEA figures found._\n")
} else {
  overview <- gsea_pdfs[!grepl("running|single_term", gsea_pdfs, ignore.case = TRUE)]
  if (!length(overview)) overview <- gsea_pdfs
  for (f in utils::head(overview, 10)) {
    cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
    cat("\nThis visualization summarizes a ranked-list enrichment result; it does not by itself demonstrate pathway activity or cellular function.\n")
  }
}
```

```{r gsea-running, results='asis'}
running <- gsea_pdfs[grepl("running|single_term", gsea_pdfs, ignore.case = TRUE)]
if (length(running)) {
  cat("<details><summary>Selected GSEA running curves</summary>\n\n")
  for (f in utils::head(running, 12)) {
    cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
  }
  cat("</details>\n")
}
```

# Custom gene sets are a separate exploratory evidence layer

Custom GSVA/ssGSEA scores should be versioned, sourced and separated from
database enrichment. A significant transcript score supports a difference in
the defined expression panel; it does not establish functional flux, pathway
activation or a causal mechanism.

```{r gsva, results='asis'}
gsva_pdfs <- list_pdfs(viz_dir, pattern = "(GSVA|CustomGeneSets).*\\.pdf$")
if (!length(gsva_pdfs)) cat("_No custom gene-set figures found or custom scoring was not requested._\n")
for (f in utils::head(gsva_pdfs, 12)) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
  cat("\nInterpret sample-level scores with effect size, multiplicity correction, gene-set provenance and overlap audit.\n")
}
```

# Cross-contrast structure helps separate shared and treatment-specific signals

```{r cross-contrast, results='asis'}
cross_pdfs <- list_pdfs(viz_dir, pattern = "(CompareCluster|overlap|UpSet|Jaccard).*\\.pdf$")
if (!length(cross_pdfs)) cat("_No cross-contrast figures found or only one contrast was analyzed._\n")
for (f in utils::head(cross_pdfs, 8)) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
  cat("\nOverlap is descriptive unless a formal equivalence, interaction or differential-effect test was prespecified.\n")
}
```

# Evidence-linked synthesis and hypotheses

```{r claims}
if (!is.null(claim_ledger) && nrow(claim_ledger)) {
  public_cols <- intersect(c("claim_id", "claim", "claim_level", "scope", "evidence", "next_evidence"), colnames(claim_ledger))
  show_table(claim_ledger[, public_cols, drop = FALSE],
             caption = "Claim-evidence ledger: conclusions remain bounded by the listed evidence level.", max_rows = 100)
} else {
  cat('<div class="caution-box"><strong>No validated claim-evidence ledger was supplied.</strong> The report therefore stops at statistical description and does not generate project-specific mechanism claims. Add a reviewed <code>claim_evidence_ledger.csv</code> to include alternatives, falsifiers and prioritized validation.</div>')
}
```

# Limitations, uncertainty, and robustness

- Small groups limit precision and outlier robustness; show individual samples and effect sizes.
- Unrecorded batch, culture, passage, dose or time variables can remain confounded with condition.
- “Not significant” does not establish equivalence or absence of a biologically meaningful effect.
- Transcript-level pathway scores do not establish protein activity, localization, secretion or functional flux.
- ORA and GSEA term names require leading-edge or contributing-gene review before biological extrapolation.

# Report coverage and unresolved questions

The report coverage manifest is the audit trail connecting all analysis domains
to the main body, a compact summary, an appendix, an explicit not-applicable
decision or a documented omission. A missing optional analysis remains visible.

```{r coverage}
show_table(coverage, caption = "Analysis-domain coverage contract.", max_rows = 100)
```

# Reproducibility and methods handoff

```{r session-info, results='asis'}
sess <- file.path(outdir, "sessionInfo.txt")
if (file.exists(sess)) {
  cat("```\n")
  cat(utils::head(readLines(sess, warn = FALSE), 60), sep = "\n")
  cat("\n```\n")
} else {
  cat("_sessionInfo.txt not found._\n")
}
```

# Complete output index {.appendix}

The main narrative intentionally avoids embedding every redundant threshold or
plot variant. This index provides exhaustive discoverability without turning
the report into an unreadable figure dump.

```{r inventory}
if (!is.null(inventory) && nrow(inventory)) {
  by_ext <- as.data.frame(table(inventory$extension), stringsAsFactors = FALSE)
  colnames(by_ext) <- c("Extension", "Files")
  show_table(by_ext, caption = "Saved output inventory by file type.")
  cat("<details><summary>All saved output paths</summary>\n\n")
  show_table(inventory, max_rows = 500)
  cat("</details>\n")
}
```
