---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    embed-resources: true
    fig-cap-location: bottom
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
params:
  title: "RNA-seq Analysis Report"
  author: ""
  outdir: "."
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
outdir <- params$outdir
deg_dir <- file.path(outdir, "1-DEG")
gsea_dir <- file.path(outdir, "2-GSEA")
viz_dir <- file.path(outdir, "3-Visualization")

read_csv_safe <- function(path, ...) {
  if (!file.exists(path)) return(NULL)
  tryCatch(utils::read.csv(path, stringsAsFactors = FALSE, ...), error = function(e) NULL)
}
list_pdfs <- function(dir, pattern = "\\.pdf$") {
  if (!dir.exists(dir)) return(character(0))
  sort(list.files(dir, pattern = pattern, full.names = FALSE, recursive = TRUE))
}
show_table <- function(df, caption = NULL, max_rows = 50) {
  if (is.null(df) || nrow(df) == 0) {
    cat("_No data available._\n")
    return(invisible(NULL))
  }
  if (nrow(df) > max_rows) {
    cat(sprintf("_Showing first %d of %d rows._\n", max_rows, nrow(df)))
    df <- utils::head(df, max_rows)
  }
  knitr::kable(df, caption = caption, row.names = FALSE)
}
show_pdf <- function(rel_path, width = "100%") {
  # Embed a PDF figure inline so it renders inside the HTML report.
  full <- file.path(viz_dir, rel_path)
  if (!file.exists(full)) {
    cat("_Figure not found: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  uri <- knitr::image_uri(full)
  cat(sprintf('<object data="%s" type="application/pdf" width="%s" height="620px">', uri, width))
  cat(sprintf('<p>PDF figure: <a href="%s">%s</a></p>', uri, rel_path))
  cat('</object>')
}
```

# Overview

This report summarizes a bulk RNA-seq analysis run with the **RNAseq-Templates**
workflow. All figures are saved as editable PDFs under `3-Visualization/`; the
sections below embed the key tables and figures for a single shareable document.

**Analysis directory:** `r outdir`

```{r summary}
summary_txt <- file.path(outdir, "Analysis_summary.txt")
if (file.exists(summary_txt)) {
  cat("```\n")
  cat(readLines(summary_txt, warn = FALSE), sep = "\n")
  cat("\n```\n")
}
```

# Differential Expression

## DEG counts by threshold

```{r deg-summary}
deg_summary <- read_csv_safe(file.path(deg_dir, "DEG_threshold_summary.csv"))
show_table(deg_summary, caption = "Number of differentially expressed genes per comparison and threshold.")
```

## Raw vs shrunken log2FC strategy

```{r lfc-strategy}
lfc_strategy <- read_csv_safe(file.path(deg_dir, "DEG_lfc_strategy_summary.csv"))
show_table(lfc_strategy, caption = "DEG counts using raw vs shrunken log2 fold-change.")
```

## Volcano plots

```{r volcano, results='asis'}
volcano_pdfs <- list_pdfs(viz_dir, pattern = "^Volcano_.*\\.pdf$")
if (length(volcano_pdfs) == 0) cat("_No volcano plots found._\n")
for (f in volcano_pdfs) {
  cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
}
```

# Sample Quality Control

```{r qc, results='asis'}
qc_pdfs <- list_pdfs(viz_dir, pattern = "^(PCA_plot|Sample_distance_heatmap).*\\.pdf$")
for (f in qc_pdfs) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
}
if (length(qc_pdfs) == 0) cat("_No QC figures found._\n")
```

# Enrichment (ORA)

```{r ora-summary}
ora_summary <- read_csv_safe(file.path(gsea_dir, "ORA_threshold_summary.csv"))
show_table(ora_summary, caption = "Over-representation analysis summary across thresholds.")
```

```{r ora-figures, results='asis'}
ora_pdfs <- list_pdfs(viz_dir, pattern = "(ORA|GO_|KEGG_).*\\.pdf$")
if (length(ora_pdfs) > 0) {
  cat("\n## ORA figures\n\n")
  for (f in utils::head(ora_pdfs, 12)) {
    cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
  }
}
```

# Enrichment (GSEA)

```{r gsea-figures, results='asis'}
gsea_pdfs <- list_pdfs(viz_dir, pattern = "GSEA.*\\.pdf$")
if (length(gsea_pdfs) == 0) {
  cat("_No GSEA figures found._\n")
} else {
  for (f in utils::head(gsea_pdfs, 12)) {
    cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
  }
}
```

# Expression Heatmaps

```{r heatmaps, results='asis'}
hm_pdfs <- list_pdfs(viz_dir, pattern = "Heatmap.*\\.pdf$")
if (length(hm_pdfs) == 0) cat("_No heatmaps found._\n")
for (f in utils::head(hm_pdfs, 8)) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
}
```

# Reproducibility

```{r session-info, results='asis'}
sess <- file.path(outdir, "sessionInfo.txt")
if (file.exists(sess)) {
  cat("```\n")
  cat(utils::head(readLines(sess, warn = FALSE), 60), sep = "\n")
  cat("\n```\n")
} else {
  cat("_sessionInfo.txt not found._\n")
}
```
