---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: cosmo
    embed-resources: true
    fig-cap-location: bottom
    df-print: kable
execute:
  echo: false
  warning: false
  message: false
params:
  title: "RNA-seq Analysis Report"
  author: ""
  outdir: "."
  primary_threshold: "standard"
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
outdir <- params$outdir
deg_dir <- file.path(outdir, "1-DEG")
gsea_dir <- file.path(outdir, "2-GSEA")
viz_dir <- file.path(outdir, "3-Visualization")

read_csv_safe <- function(path, ...) {
  if (!file.exists(path)) return(NULL)
  tryCatch(utils::read.csv(path, stringsAsFactors = FALSE, ...), error = function(e) NULL)
}
list_pdfs <- function(dir, pattern = "\\.pdf$") {
  if (!dir.exists(dir)) return(character(0))
  sort(list.files(dir, pattern = pattern, full.names = FALSE, recursive = TRUE))
}
show_table <- function(df, caption = NULL, max_rows = 50) {
  if (is.null(df) || nrow(df) == 0) {
    cat("_No data available._\n")
    return(invisible(NULL))
  }
  if (nrow(df) > max_rows) {
    cat(sprintf("_Showing first %d of %d rows._\n", max_rows, nrow(df)))
    df <- utils::head(df, max_rows)
  }
  knitr::kable(df, caption = caption, row.names = FALSE)
}
show_pdf <- function(rel_path, width = "100%") {
  # Render the first PDF page to PNG before embedding. Browser PDF objects are
  # inconsistent and can turn an otherwise valid figure into a blank panel.
  full <- file.path(viz_dir, rel_path)
  if (!file.exists(full)) {
    cat("_Figure not found: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  if (!nzchar(Sys.which("pdftoppm"))) {
    cat("_PDF preview unavailable because `pdftoppm` is not installed: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  base <- tempfile(pattern = "rnaseq-report-")
  png <- paste0(base, ".png")
  code <- suppressWarnings(system2(
    "pdftoppm",
    c("-f", "1", "-l", "1", "-singlefile", "-r", "150", "-png",
      shQuote(full), shQuote(base)),
    stdout = FALSE, stderr = FALSE
  ))
  if (!identical(code, 0L) || !file.exists(png)) {
    cat("_PDF preview failed: ", rel_path, "_\n", sep = "")
    return(invisible(NULL))
  }
  uri <- knitr::image_uri(png)
  cat(sprintf(
    '<figure><img src="%s" alt="%s" style="width:%s;height:auto"><figcaption>%s</figcaption></figure>',
    uri, rel_path, width, rel_path
  ))
}

coverage <- read_csv_safe(file.path(outdir, "report_coverage_manifest.csv"))
claim_ledger <- read_csv_safe(file.path(outdir, "claim_evidence_ledger.csv"), check.names = FALSE)
review_checklist <- read_csv_safe(file.path(outdir, "report_review_checklist.csv"), check.names = FALSE)
inventory <- if (dir.exists(outdir)) {
  paths <- list.files(outdir, recursive = TRUE, full.names = FALSE)
  data.frame(path = paths, extension = tolower(tools::file_ext(paths)), stringsAsFactors = FALSE)
} else NULL

# Render a per-section project annotation from report_interpretation/<key>.md.
# Whole-line HTML comments count as scaffold guidance and are stripped; a file
# with no remaining content renders nothing, so untouched scaffold files stay
# hidden. Annotations knit with the project environment so inline `r ` code
# works, and they persist across re-renders without editing this template.
show_interpretation <- function(key) {
  path <- file.path(outdir, "report_interpretation", paste0(key, ".md"))
  if (!file.exists(path)) return(invisible(NULL))
  txt <- readLines(path, warn = FALSE)
  txt <- txt[!grepl("^\\s*<!--.*-->\\s*$", txt)]
  if (!any(nzchar(trimws(txt)))) return(invisible(NULL))
  cat("\n::: {.interpretation-box}\n\n")
  cat("*Project annotation — edit `report_interpretation/", key, ".md`*\n\n", sep = "")
  cat(knitr::knit(text = txt, quiet = TRUE))
  cat("\n:::\n\n")
}

# --- Data-driven result-detail helpers ---------------------------------------
# These surface the actual genes and enriched terms behind each result block so
# the report answers "which genes / which pathways" directly, for any project.
# Every helper degrades to NULL when the underlying file is absent, so the
# report still renders for partial runs.

# Primary DEG threshold: honor the run's declared default when that threshold
# has result tables; retain the standard/first-folder fallback for old bundles.
primary_threshold <- function() {
  th <- list.dirs(deg_dir, recursive = FALSE, full.names = FALSE)
  th <- th[vapply(th, function(x) {
    length(list.files(file.path(deg_dir, x), pattern = "^DEG_results_.*\\.csv$")) > 0
  }, logical(1))]
  declared <- params$primary_threshold
  if (is.character(declared) && length(declared) == 1 && declared %in% th) return(declared)
  if ("standard" %in% th) "standard" else if (length(th)) th[[1]] else NULL
}

# Contrast names discovered from the primary-threshold DEG result files.
list_contrasts <- function(threshold = primary_threshold()) {
  if (is.null(threshold)) return(character(0))
  files <- list.files(file.path(deg_dir, threshold), pattern = "^DEG_results_.*\\.csv$")
  sub("^DEG_results_(.*)\\.csv$", "\\1", files)
}

# Top `n` up- and down-regulated genes per contrast by |shrunken log2FC|.
top_deg_table <- function(contrast, threshold = primary_threshold(), n = 8) {
  df <- read_csv_safe(file.path(deg_dir, threshold, paste0("DEG_results_", contrast, ".csv")))
  if (is.null(df) || !nrow(df) || !"gene_name" %in% colnames(df)) return(NULL)
  lfc_col <- intersect(c("log2FoldChange_shrunken", "log2FoldChange", "log2FoldChange_raw"), colnames(df))[1]
  if (is.na(lfc_col)) return(NULL)
  df <- df[!is.na(df[[lfc_col]]), , drop = FALSE]
  pick <- function(x, dirn) {
    if (is.null(x) || !nrow(x)) return(NULL)
    x <- utils::head(x[order(-abs(x[[lfc_col]])), , drop = FALSE], n)
    data.frame(
      Direction = dirn, Gene = x$gene_name,
      log2FC = round(x[[lfc_col]], 2),
      padj = if ("padj" %in% colnames(x)) signif(x$padj, 3) else NA_real_,
      stringsAsFactors = FALSE
    )
  }
  if ("significance" %in% colnames(df)) {
    out <- rbind(pick(df[grepl("^up", df$significance, ignore.case = TRUE), , drop = FALSE], "Up"),
                 pick(df[grepl("^down", df$significance, ignore.case = TRUE), , drop = FALSE], "Down"))
  } else {
    out <- rbind(pick(df[df[[lfc_col]] > 0, , drop = FALSE], "Up"),
                 pick(df[df[[lfc_col]] < 0, , drop = FALSE], "Down"))
  }
  if (is.null(out) || !nrow(out)) NULL else out
}

# Top enriched ORA terms for one contrast, merging the available
# direction/ontology source tables saved under the primary threshold.
ora_top_terms <- function(contrast, threshold = primary_threshold(), n = 6) {
  dir_path <- file.path(gsea_dir, threshold)
  if (!dir.exists(dir_path)) return(NULL)
  files <- list.files(dir_path, pattern = paste0("ORA.*", contrast, "\\.csv$"), full.names = TRUE)
  if (!length(files)) return(NULL)
  rows <- lapply(files, function(f) {
    df <- read_csv_safe(f)
    if (is.null(df) || !nrow(df) || !"Description" %in% colnames(df)) return(NULL)
    db <- if (grepl("KEGG", basename(f))) "KEGG" else "GO"
    dirn <- if (grepl("_UP_", basename(f))) "Up" else if (grepl("_DOWN_", basename(f))) "Down" else "Both"
    df <- utils::head(df[order(df$p.adjust), , drop = FALSE], n)
    data.frame(Ontology = db, Direction = dirn, Term = df$Description,
               Count = if ("Count" %in% colnames(df)) df$Count else NA_integer_,
               padj = signif(df$p.adjust, 3), stringsAsFactors = FALSE)
  })
  out <- do.call(rbind, rows)
  if (is.null(out) || !nrow(out)) NULL else out[order(out$padj), , drop = FALSE]
}

# Top GSEA gene sets for one contrast, split by enrichment direction (NES sign).
gsea_top_terms <- function(contrast, n = 5) {
  files <- c(GO = file.path(gsea_dir, paste0("GSEA_GO_", contrast, ".csv")),
             KEGG = file.path(gsea_dir, paste0("GSEA_KEGG_", contrast, ".csv")))
  rows <- lapply(names(files), function(db) {
    df <- read_csv_safe(files[[db]])
    if (is.null(df) || !nrow(df) || !all(c("Description", "NES") %in% colnames(df))) return(NULL)
    df <- df[!is.na(df$NES) & !is.na(df$p.adjust), , drop = FALSE]
    if (!nrow(df)) return(NULL)
    pick <- function(x, dirn) {
      if (!nrow(x)) return(NULL)
      x <- utils::head(x[order(x$p.adjust), , drop = FALSE], n)
      data.frame(Ontology = db, Direction = dirn, Term = x$Description,
                 NES = round(x$NES, 2), padj = signif(x$p.adjust, 3), stringsAsFactors = FALSE)
    }
    rbind(pick(df[df$NES > 0, , drop = FALSE], "Up"), pick(df[df$NES < 0, , drop = FALSE], "Down"))
  })
  out <- do.call(rbind, rows)
  if (is.null(out) || !nrow(out)) NULL else out[order(out$padj), , drop = FALSE]
}
```

<style>
.summary-box {background:#f3f7fb;border-left:5px solid #3778a8;padding:0.9rem 1.1rem;margin:1rem 0;}
.caution-box {background:#fff8e8;border-left:5px solid #d99a2b;padding:0.9rem 1.1rem;margin:1rem 0;}
.interpretation-box {background:#f6f3fb;border-left:5px solid #7a5fa8;padding:0.7rem 1.1rem;margin:1rem 0;}
.interpretation-box > p:first-of-type em {color:#7a5fa8;font-size:0.85rem;}
figure {margin:1.2rem 0 1.6rem 0;}
figure img {border:1px solid #e5e7eb;}
figcaption {color:#5f6368;font-size:0.9rem;margin-top:0.45rem;}
details {margin:0.8rem 0;}
</style>

# Technical summary

<div class="summary-box">

This report summarizes the saved outputs of an **RNAseq-Templates** run. It is
designed as an evidence-linked technical report: primary results appear in the
main reading path, sensitivity and exhaustive output indexes appear later, and
missing analysis domains remain visible in the coverage table.

Project-specific biological mechanisms are reported only when a validated
`claim_evidence_ledger.csv` is supplied. In its absence, this report describes
the observed statistical results without inventing a mechanism narrative.

</div>

**Analysis directory:** `r outdir`

```{r summary}
summary_txt <- file.path(outdir, "Analysis_summary.txt")
if (file.exists(summary_txt)) {
  cat("```\n")
  cat(readLines(summary_txt, warn = FALSE), sep = "\n")
  cat("\n```\n")
}
```

```{r interpretation, results='asis'}
# Optional project-supplied biological interpretation. A project can drop a
# `report_interpretation.md` into the output directory to add a reviewed,
# project-specific narrative (top-DEG functional reading, pathway meaning,
# gene-set rationale, links to the study question) that persists across
# re-renders without editing this shared template. Finer-grained annotations
# belong in `report_interpretation/<section>.md`, rendered in place at the end
# of each matching section (see `show_interpretation()` calls below).
interp_md <- file.path(outdir, "report_interpretation.md")
if (file.exists(interp_md)) {
  cat("\n# Biological interpretation\n\n")
  cat('<div class="summary-box">Project-supplied, reviewed interpretation. It is subordinate to the statistical evidence in the sections that follow and does not replace the claim-evidence ledger.</div>\n\n')
  cat(knitr::knit(text = readLines(interp_md, warn = FALSE), quiet = TRUE))
  cat("\n")
}
```

# Scope, samples, and inferential definitions

The experimental unit, inclusion decisions, group sizes, design formula,
contrasts, primary threshold and multiple-testing families should be defined
before interpreting differential expression or enrichment. The saved metadata
below is the report's source of truth for sample inclusion.

```{r sample-design}
coldata <- read_csv_safe(file.path(deg_dir, "colData.csv"), check.names = FALSE)
show_table(coldata, caption = "Samples included in the analysis and their modeled groups.", max_rows = 100)
```

```{r interpretation-scope, results='asis'}
show_interpretation("scope")
```

# Sample quality supports—or limits—the modeled comparisons

QC figures are used to identify gross data-quality problems, unexpected sample
structure and possible outliers. Visual separation is not evidence of a
mechanism, and samples should not be removed simply because they weaken the
expected result.

```{r qc, results='asis'}
qc_pdfs <- list_pdfs(viz_dir, pattern = "^(PCA_plot|Sample_distance_heatmap|Sample_correlation|Raw_count_distribution|Filter_retention|Sample_QC_metrics).*\\.pdf$")
for (f in utils::head(qc_pdfs, 8)) {
  cat("\n## ", sub("\\.pdf$", "", basename(f)), "\n\n")
  show_pdf(f)
  cat("\nRead this figure at the sample level; any exclusion requires a predeclared QC rule or an independently documented technical failure.\n")
}
if (length(qc_pdfs) == 0) cat("_No QC figures found._\n")
```

```{r interpretation-qc, results='asis'}
show_interpretation("qc")
```

# Differential expression is reported at one primary threshold

## DEG counts by threshold

```{r deg-summary}
deg_summary <- read_csv_safe(file.path(deg_dir, "DEG_threshold_summary.csv"))
show_table(deg_summary, caption = "Number of differentially expressed genes per comparison and threshold.")
```

```{r declared-primary-threshold, results='asis'}
primary_th <- primary_threshold()
if (!is.null(primary_th)) {
  cat("The declared primary threshold for the detailed tables below is <code>", primary_th, "</code>.\n\n", sep = "")
}
```

```{r exploratory-tier-note, results='asis'}
# Nominal-p tiers are hypothesis-generation only; flag them explicitly wherever
# the threshold summary is shown so they are never read as inferential results.
if (!is.null(deg_summary) && "pvalue_column" %in% colnames(deg_summary) &&
    any(deg_summary$pvalue_column == "pvalue", na.rm = TRUE)) {
  cat('\n<div class="caution-box"><strong>Exploratory nominal-P tier present.</strong> ',
      'Rows with <code>pvalue_column = pvalue</code> use unadjusted P values to keep ',
      'pathway analysis alive in low-DEG projects. They are hypothesis-generation inputs: ',
      'ORA or gene lists derived from them require confirmation at an adjusted-P tier or by ',
      'rank-based (GSEA/GSVA) evidence before supporting any claim beyond E0/E1.</div>\n')
}
```

The declared primary threshold should drive headline counts. Other rows are
sensitivity analyses and must not be searched post hoc for the most favorable
result. DEG counts depend on effect magnitude, dispersion, expression level and
sample size; they are not a standalone measure of treatment strength.

## Raw vs shrunken log2FC strategy

```{r lfc-strategy}
lfc_strategy <- read_csv_safe(file.path(deg_dir, "DEG_lfc_strategy_summary.csv"))
show_table(lfc_strategy, caption = "DEG counts using raw vs shrunken log2 fold-change.")
```

## Top differentially expressed genes

The genes driving each headline comparison, ranked by shrunken effect size at
the primary threshold. These identify *which* genes change before any
functional interpretation is layered on.

```{r top-degs, results='asis'}
contrasts_deg <- list_contrasts()
if (!length(contrasts_deg)) {
  cat("_No per-contrast DEG tables found at the primary threshold._\n")
} else {
  for (cc in contrasts_deg) {
    tt <- top_deg_table(cc)
    if (is.null(tt)) next
    cat("\n### ", cc, "\n\n")
    print(knitr::kable(tt, caption = paste("Top up- and down-regulated genes by |shrunken log2FC|:", cc), row.names = FALSE))
    cat("\n")
  }
}
```

## Volcano plots

```{r volcano, results='asis'}
volcano_pdfs <- list_pdfs(viz_dir, pattern = "^Volcano_.*\\.pdf$")
if (length(volcano_pdfs) == 0) cat("_No volcano plots found._\n")
for (f in volcano_pdfs) {
  cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
}
```

## Expression heatmaps and selected genes

```{r heatmaps, results='asis'}
hm_pdfs <- list_pdfs(viz_dir, pattern = "(Heatmap_topDEG|Heatmap_key_genes|DEG_barplot_).*\\.pdf$")
if (length(hm_pdfs) == 0) cat("_No DEG or key-gene heatmaps/plots found._\n")
for (f in utils::head(hm_pdfs, 12)) {
  cat("\n### ", sub("\\.pdf$", "", basename(f)), "\n\n")
  show_pdf(f)
  cat("\nSelected-gene views support candidate inspection but do not create an independent statistical test.\n")
}
```

```{r interpretation-deg, results='asis'}
show_interpretation("deg")
```

# ORA findings depend on the DEG threshold and tested background

```{r ora-summary}
ora_summary <- read_csv_safe(file.path(gsea_dir, "ORA_threshold_summary.csv"))
show_table(ora_summary, caption = "Over-representation analysis summary across thresholds.")
```

ORA asks whether a thresholded DEG list contains more genes from a term than
expected against the declared expressed-gene background. Direction, threshold,
ontology and correction family must be retained when interpreting a term.

## Top enriched ORA terms

The strongest enriched GO/KEGG terms per comparison at the primary threshold,
merged across the direction-specific source tables. Disease-labelled KEGG terms
(e.g. viral infections) usually reflect shared interferon / antigen / cytotoxic
gene content rather than literal infection.

```{r ora-top-terms, results='asis'}
contrasts_ora <- list_contrasts()
if (!length(contrasts_ora)) {
  cat("_No per-contrast ORA tables found at the primary threshold._\n")
} else {
  for (cc in contrasts_ora) {
    tt <- ora_top_terms(cc)
    if (is.null(tt)) next
    cat("\n### ", cc, "\n\n")
    print(knitr::kable(tt, caption = paste("Top enriched ORA terms (GO/KEGG, by adjusted P):", cc), row.names = FALSE))
    cat("\n")
  }
}
```

```{r ora-figures, results='asis'}
ora_pdfs <- list_pdfs(viz_dir, pattern = "(ORA|GO_|KEGG_).*\\.pdf$")
if (length(ora_pdfs) > 0) {
  cat("\n## Representative ORA evidence\n\n")
  preferred <- ora_pdfs[grepl("(^|/)standard/", ora_pdfs)]
  if (!length(preferred)) preferred <- ora_pdfs
  for (f in utils::head(preferred, 10)) {
    cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
    cat("\nInterpret the displayed term together with its gene count, direction, adjusted P value and threshold-specific source table.\n")
  }
} else {
  cat("_No ORA figures found or ORA was not applicable._\n")
}
```

```{r interpretation-ora, results='asis'}
show_interpretation("ora")
```

# GSEA provides rank-based program evidence, not literal pathway function

GSEA uses a full ranked gene list and therefore reduces dependence on a single
DEG cutoff. GO terms can be redundant and disease-labelled KEGG terms often
reflect shared ribosomal, mitochondrial or inflammatory genes. Leading-edge
genes—not the pathway title alone—should anchor biological interpretation.

## Top enriched gene sets by NES

The strongest positively and negatively enriched gene sets per comparison,
ranked by adjusted P value. Positive NES = enriched among up-regulated genes;
negative NES = enriched among down-regulated genes.

```{r gsea-top-terms, results='asis'}
contrasts_gsea <- list_contrasts()
if (!length(contrasts_gsea)) {
  cat("_No per-contrast GSEA tables found._\n")
} else {
  for (cc in contrasts_gsea) {
    tt <- gsea_top_terms(cc)
    if (is.null(tt)) next
    cat("\n### ", cc, "\n\n")
    print(knitr::kable(tt, caption = paste("Top GSEA gene sets by direction (GO/KEGG):", cc), row.names = FALSE))
    cat("\n")
  }
}
```

```{r gsea-figures, results='asis'}
gsea_pdfs <- list_pdfs(viz_dir, pattern = "GSEA.*\\.pdf$")
if (length(gsea_pdfs) == 0) {
  cat("_No GSEA figures found._\n")
} else {
  overview <- gsea_pdfs[!grepl("running|single_term", gsea_pdfs, ignore.case = TRUE)]
  if (!length(overview)) overview <- gsea_pdfs
  for (f in utils::head(overview, 10)) {
    cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
    cat("\nThis visualization summarizes a ranked-list enrichment result; it does not by itself demonstrate pathway activity or cellular function.\n")
  }
}
```

```{r gsea-running, results='asis'}
running <- gsea_pdfs[grepl("running|single_term", gsea_pdfs, ignore.case = TRUE)]
if (length(running)) {
  cat("<details><summary>Selected GSEA running curves</summary>\n\n")
  for (f in utils::head(running, 12)) {
    cat("\n### ", sub("\\.pdf$", "", f), "\n\n")
    show_pdf(f)
  }
  cat("</details>\n")
}
```

```{r interpretation-gsea, results='asis'}
show_interpretation("gsea")
```

# Custom gene sets are a separate exploratory evidence layer

Custom GSVA/ssGSEA scores should be versioned, sourced and separated from
database enrichment. A significant transcript score supports a difference in
the defined expression panel; it does not establish functional flux, pathway
activation or a causal mechanism.

```{r gsva, results='asis'}
gsva_pdfs <- list_pdfs(viz_dir, pattern = "(GSVA|CustomGeneSets).*\\.pdf$")
if (!length(gsva_pdfs)) cat("_No custom gene-set figures found or custom scoring was not requested._\n")
for (f in utils::head(gsva_pdfs, 12)) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
  cat("\nInterpret sample-level scores with effect size, multiplicity correction, gene-set provenance and overlap audit.\n")
}
```

```{r interpretation-custom-genesets, results='asis'}
show_interpretation("custom_genesets")
```

# Cross-contrast structure helps separate shared and treatment-specific signals

```{r cross-contrast, results='asis'}
cross_pdfs <- list_pdfs(viz_dir, pattern = "(CompareCluster|overlap|UpSet|Jaccard).*\\.pdf$")
if (!length(cross_pdfs)) cat("_No cross-contrast figures found or only one contrast was analyzed._\n")
for (f in utils::head(cross_pdfs, 8)) {
  cat("\n## ", sub("\\.pdf$", "", f), "\n\n")
  show_pdf(f)
  cat("\nOverlap is descriptive unless a formal equivalence, interaction or differential-effect test was prespecified.\n")
}
```

```{r interpretation-cross-contrast, results='asis'}
show_interpretation("cross_contrast")
```

# Evidence-linked synthesis and hypotheses

```{r claims}
if (!is.null(claim_ledger) && nrow(claim_ledger)) {
  public_cols <- intersect(c("claim_id", "claim", "claim_level", "scope", "evidence", "next_evidence"), colnames(claim_ledger))
  show_table(claim_ledger[, public_cols, drop = FALSE],
             caption = "Claim-evidence ledger: conclusions remain bounded by the listed evidence level.", max_rows = 100)
} else {
  cat('<div class="caution-box"><strong>No validated claim-evidence ledger was supplied.</strong> The report therefore stops at statistical description and does not generate project-specific mechanism claims. Add a reviewed <code>claim_evidence_ledger.csv</code> to include alternatives, falsifiers and prioritized validation.</div>')
}
```

```{r interpretation-synthesis, results='asis'}
show_interpretation("synthesis")
```

# Limitations, uncertainty, and robustness

- Small groups limit precision and outlier robustness; show individual samples and effect sizes.
- Unrecorded batch, culture, passage, dose or time variables can remain confounded with condition.
- “Not significant” does not establish equivalence or absence of a biologically meaningful effect.
- Transcript-level pathway scores do not establish protein activity, localization, secretion or functional flux.
- ORA and GSEA term names require leading-edge or contributing-gene review before biological extrapolation.

```{r interpretation-limitations, results='asis'}
show_interpretation("limitations")
```

# Report coverage and unresolved questions

The report coverage manifest is the audit trail connecting all analysis domains
to the main body, a compact summary, an appendix, an explicit not-applicable
decision or a documented omission. A missing optional analysis remains visible.

```{r coverage}
show_table(coverage, caption = "Analysis-domain coverage contract.", max_rows = 100)
```

## Scientific review status

Draft rendering and scientific approval are separate. A report is publishable
only after all checklist rows are approved or explicitly marked not applicable
with a reason.

```{r review-checklist}
show_table(review_checklist, caption = "Human scientific-review checklist.", max_rows = 100)
```

# Reproducibility and methods handoff

```{r session-info, results='asis'}
sess <- file.path(outdir, "sessionInfo.txt")
if (file.exists(sess)) {
  cat("```\n")
  cat(utils::head(readLines(sess, warn = FALSE), 60), sep = "\n")
  cat("\n```\n")
} else {
  cat("_sessionInfo.txt not found._\n")
}
```

# Complete output index {.appendix}

The main narrative intentionally avoids embedding every redundant threshold or
plot variant. This index provides exhaustive discoverability without turning
the report into an unreadable figure dump.

```{r inventory}
if (!is.null(inventory) && nrow(inventory)) {
  by_ext <- as.data.frame(table(inventory$extension), stringsAsFactors = FALSE)
  colnames(by_ext) <- c("Extension", "Files")
  show_table(by_ext, caption = "Saved output inventory by file type.")
  cat("<details><summary>All saved output paths</summary>\n\n")
  show_table(inventory, max_rows = 500)
  cat("</details>\n")
}
```
